### 2022-06-23 manuscript.tex
Merge pull request #241 from SORTEE-Github-Hackathon/phpb-conciseness-03.use-cases-b.md- [ci skip]
This build is based on
https://github.com/SORTEE-Github-Hackathon/manuscript/commit/dc28d7dd993d0f5ca61b1bd7fc59d1cc22e39ce2.

This commit was created by the following CI build and job:
https://github.com/SORTEE-Github-Hackathon/manuscript/commit/dc28d7dd993d0f5ca61b1bd7fc59d1cc22e39ce2/checks
https://github.com/SORTEE-Github-Hackathon/manuscript/actions/runs/2550899642

This manuscript
(\href{https://SORTEE-Github-Hackathon.github.io/manuscript/v/4ac5f7eeaa29350c0b737ed7e3e45fa0be45da12/}{permalink})
was automatically generated
from \href{https://github.com/SORTEE-Github-Hackathon/manuscript/tree/4ac5f7eeaa29350c0b737ed7e3e45fa0be45da12}{SORTEE-Github-Hackathon/manuscript@4ac5f7e}
on June 15, 2022.

@@

\hypertarget{creating-a-website}{%
\subsubsection{Creating a website}\label{creating-a-website}}

Personal or lab websites are one method of improving the dissemination of research findings and coordination of research efforts, but most EEB researchers have little experience in building or hosting webpages.
GitHub Pages (\url{https://pages.github.com}) is a feature that, when activated for a repository, renders content written in markdown (e.g., \url{https://github.com/SORTEE-Github-Hackathon/main-website}) as an HTML web page with a URL (e.g., \url{https://sortee-github-hackathon.github.io/main-website/}).
This can be useful for sharing reports (e.g.~written with RMarkdown and rendered to HTML) with collaborators, or for more complex projects like personal, project, or lab websites.
GitHub also offers website templates and it is possible to create a website by forking (\protect\hyperlink{definitions}{Box 1}) a repository hosting a GitHub Pages site as a starting point.
Aside from free hosting services, another benefit is that GitHub Pages are autogenerated, meaning that when content is modified in the associated GitHub repository, the website instantly updates\textsuperscript{\protect\hyperlink{ref-RVetqmsg}{7}}.
Though the templates are useful for quickly starting up a new website, users are able to fully customize their Pages websites (for technical details of customizing GitHub Pages site see\textsuperscript{\protect\hyperlink{ref-MXxgZJ45}{28}}).
We emphasize that despite the many benefits of using GitHub pages (free hosting, templates, customization), this avenue for creating a website will often be more time intensive than the out-of-the-box platforms. ``Static websites'' (i.e., websites where every user sees the same content) can be created using Jekyll (\url{https://jekyllrb.com}) and Hugo (\url{https://gohugo.io}), which also include template libraries for websites that can be hosted freely via GitHub pages.
Several R packages such as distill (\url{https://github.com/rstudio/distill}) and blogdown (\url{https://bookdown.org/yihui/blogdown}) have made the interface to these website generators more accessible.
In addition, Quarto (\url{https://quarto.org}) has options for creating websites (\url{https://quarto.org/docs/websites}) as well as the code-free website generator wowchemy (\url{https://wowchemy.com}) which uses Hugo.

@@

GitHub makes it easy to store and share a variety of data files in the cloud.
If a repository is made ``public'' the URL to the repository can be shared freely with others.
However, for a variety of reasons (e.g., privately owned company, ability to make repositories private, accounts can be deleted at will) GitHub is not considered a long-term data or code repository like Zenodo and Figshare\textsuperscript{\protect\hyperlink{ref-10ghgV3S8}{3},\protect\hyperlink{ref-kEX5dgzK}{9},\protect\hyperlink{ref-tbl:compare}{\textbf{tbl:compare?}}}.
Also, unlike long-term repositories, GitHub does not issue Digital Object Identifiers (DOIs) for content uploaded to their servers.
DOIs are persistent and unique alpha-numeric IDs assigned to research products like papers, code, and data.
DOIs allows tracking and citing research products.
For this reason, scientists who share code and data through GitHub are strongly encouraged to also submit GitHub repository content to a long-term data archive\textsuperscript{\protect\hyperlink{ref-1Du6fzB8g}{29}}.
Fortunately, both long-term repositories mentioned above (Zenodo and Figshare) have integrations with GitHub.
After linking one's GitHub repository to Zenodo or Figshare, every time a release (\protect\hyperlink{definitions}{Box 1}) is made, a snapshot of the entire repository is archived with a versioned, citable DOI.
Linking one's GitHub repository with Zenodo, etc. to obtain a DOI helps work become findable, gives proper attribution, and can ensure long-term stability\textsuperscript{\protect\hyperlink{ref-iIEKCTLU}{30}}.

Many researchers believe that their code is not useful because their analysis is context-specific and not explicitly designed for re-use like software.
However, there are many reasons to share data and code beyond re-use.
Even if code lacks detailed annotation, it shows the exact steps taken to conduct an analysis, and therefore provides the most detailed look into how to reproduce a given analysis\textsuperscript{\protect\hyperlink{ref-uBJwnPbq}{31}}.
This is important in light of the reproducibility crisis\textsuperscript{\protect\hyperlink{ref-1HZdsK5Kn}{32}} and will become increasingly important to the collective scientific enterprise as advances in computing power and accessibility unlock the ability to conduct `big data' meta research with data that has already been collected by others.
Failing to include data and code with publications leaves future scientists with fewer open resources from which to understand the world.

An important aspect of making code citable and reusable is adding an appropriate licence.
Code without a licence is (by default) actually provided more protection from reuse than code with an open licence.
The standard GitHub licensing options are best suited for software, so depending on how you want your code to be used you can choose one of the other license options.
For example, if your code is intended only for your specific analysis, consider a Creative Commons License.
If you wish to receive attribution for any reuse of your code, consider a CC BY 4.0 license.
If you have built an app, tool, package, or other product that you would like others to use and would like attribution for any reuse of your code, consider the GNU General Public License v3.
This license also prohibits the re-user from making their re-used version private.
To help navigate through the potential licenses available to you and their attributes the Choose a License (\url{https://choosealicense.com/non-software/}) website can offer further guidance.

\hypertarget{collaborative-code-editing-and-asynchronous-working}{%
\subsubsection{Collaborative (code) editing and asynchronous working}\label{collaborative-code-editing-and-asynchronous-working}}

GitHub facilitates collaborative coding and asynchronous working because researchers can seamlessly access and contribute to data and code regardless of disparities in schedules or location.
Project work can also be clearly split among team members, giving them the flexibility to contribute when it best fits their schedule.
We acknowledge that there are important differences between the average software developer and EEB researcher using GitHub, and that not all of GitHub's features are useful for research workflows (\protect\hyperlink{tbl:roles}{Table 1}).
However, git's core features such as forking and branching (\protect\hyperlink{definitions}{Box 1}) allow for simultaneous coding on different versions of the same research project, and alternative versions can be discussed and resolved with GitHub.
When EEB projects are contained on the GitHub platform, researchers can stay abreast of progress made by other collaborators without the need for meetings or emails.
Version control\textsuperscript{\protect\hyperlink{ref-K7nbP1Ty}{33}} features also allow users to make progress and changes without worrying about irreparably writing over someone else's work.
By enabling more comprehensive remote collaboration, GitHub encourages the exchange of ideas among researchers at different institutions and in different countries, which can serve to improve the quality of the research itself by providing open access to data and code.
In academic settings, GitHub can also facilitate interactions between research advisors and advisees, providing a platform for students or other trainees to share in-progress code, and flag specific challenges or questions for their supervisors or mentors.
Periodic code review\textsuperscript{\protect\hyperlink{ref-hm9PaCLD}{34}} can also help advisors identify errors early in the process, and inform further training and mentorship to fill gaps in skills.

@@

Beyond supporting collaboration at the level of code, GitHub can be used for collaboratively writing manuscripts.
Writing a manuscript in GitHub and storing it with associated data and code all in the same repository increases scientific reproducibility because files associated with a manuscript can be found in one place.
Co-authors can contribute new text to a manuscript or suggest revisions through GitHub's robust pull request feature (\protect\hyperlink{definitions}{Box 1}) which provides a line-by-line view of all proposed changes.
Further, authors can make use of the Discussions tab to suggest relevant papers to be cited, and can raise issues during the writing process that can be assigned to collaborators (\protect\hyperlink{tbl:roles}{Table 1}).

While GitHub is not considered as user-friendly for manuscript development as conventional text processors\textsuperscript{\protect\hyperlink{ref-4ny1onB0}{8}}, it has been substantially improved with recent tools.
Manuscripts can be written on GitHub with Markdown.
Add-ons like HackMD (\url{https://hackmd.io}), can enable real-time collaboration for individual Markdown documents.
Many tools exist for extending Markdown and Pandoc to add formatting features necessary for scientific writing like in-text citations and figure and table cross references.

We wrote this manuscript using Manubot\textsuperscript{\protect\hyperlink{ref-YuJbg3zO}{35}}, a collaborative manuscript platform that uses Markdown for writing and GitHub for storing and tracking changes to a manuscript over time.
Manubot uses a GitHub Actions-based typesetting system to compile individual Markdown files stored in a GitHub repository into a single LaTeX document, which can be displayed in Word, HTML, or PDF formats.
The resulting manuscript can also be compiled using a journal's .tex template to match their formatting requirements.
Since this tool reruns the entire manuscript compilation process with any change to the underlying repository, it can also accomodate continuous integration of code updates into figures and tables with additional GitHub Actions (e.g.~https://github.com/SORTEE-Github-Hackathon/manuscript/tree/main/.github/workflows).
Manubot also allows for straightforward citation management based on URLs or DOIs.
For more examples of Manubot being used for manuscripts, see\textsuperscript{\protect\hyperlink{ref-YuJbg3zO}{35}}.

Since Manubot works on documents in a distributed format, it can be difficult to edit manuscripts for overall flow with only this tool.
We employed hypothes.is (\url{https://hypothes.is}) to write comments on the HTML manuscript document produced by Manubot, which we then addressed by committing changes to the underlying Markdown files via pull requests.
Other authors can reply to the comments to indicate agreement or disagreement, and to note when changes have been made.
Other tools can also be used for version control of scientific manuscripts including R Markdown via the bookdown package (\url{https://bookdown.org/}), jupyter notebooks\textsuperscript{\protect\hyperlink{ref-i4FOZpjN}{36}} and a relatively new tool, Quarto (\url{https://quarto.org}).

\hypertarget{peer-review}{%
\subsubsection{Peer-Review}\label{peer-review}}

Peer review plays a critical role in the scientific evaluation of EEB research.
GitHub provides an open and transparent platform that can be used for either directly providing feedback on research products or building a to-do list from reviewer comments.
Peer review of research code by the open software development community rOpenSci (\url{https://ropensci.org/software-review/}) and of research software and associated manuscripts by the Journal of Open Source Software (\url{https://joss.readthedocs.io/en/latest/submitting.html}) requires that submitted work is hosted on GitHub and their review processes make use of GitHub issues.
rOpenSci's efforts have resulted in many well-used R packages for ecology research including rfishbase\textsuperscript{\protect\hyperlink{ref-Xsdcv6q}{37}} and taxize\textsuperscript{\protect\hyperlink{ref-FVBWKkZu}{38}}.

GitHub can also be used as a hub for reviewers and authors during the peer review process of an ordinary research manuscript.
If the code associated with a manuscript is made available at the time of submission (e.g.~via a link to a GitHub repository in a Data Availability Statement), peer-reviewers may be able to offer more helpful suggestions on written methods and may even make comments on the code itself, potentially catching bugs or errors before publication.
GitHub can also be used as a hub for reviewers and authors during the peer review process of an ordinary research manuscript.
GitHub issues can be used to organize and discuss reviewer suggestions and to assign them to co-authors (See example in \url{https://github.com/BrunaLab/HeliconiaDemography/issues?q=is\%3Aissue+label\%3A\%22reviewer+comment\%22+}).
When reviewer comments are posted as separate issues, authors can comment on the issues to discuss possible changes and assign themselves to indicate which comments they intend to handle.
Co-authors can then integrate their edits and responses to reviewers using pull requests.

@@

Conventional research practices typically rely on one or two people running and checking the data analyses, while most coauthors (and readers of the subsequent publication) see only the final results and a verbal description of the analytical steps.
In the developmental stages, collaborators can directly see the code for the analysis, manipulate and explore the data themselves, and check for errors.
Cynically, there is also more insurance against nefarious colleagues that may be tempted to distort results\textsuperscript{\protect\hyperlink{ref-18PTmKJkq}{39}}.
Collaborators as well as pre-publication and post-publication reviewers are better positioned to discover questionable findings if they have full and transparent access to the project.

This transparency can similarly be extended beyond coauthors to the entire scientific community.
Publishing the data and reproducible workflows along with the manuscript allows any reader to review the analysis and reproduce the experiment\textsuperscript{\protect\hyperlink{ref-NOgBWVAr}{40}}.
Supplying code for (novel) methods that are proposed or used also reduces barriers to knowledge and can greatly improve the ability of others to build on existing work, resulting in greater proliferation and accessibility for a broader audience.
GitHub even provides a useful Discussions Forum (\url{https://docs.github.com/en/discussions}) that aids the direct communication with repository owners, as well as the GitHub Community (\url{https://github.community/}) forum for more general questions and sharing of expertise (\protect\hyperlink{tbl:roles}{Table 1}.

@@

The development of research code continues, and so does the need to consider project continuity, especially in EEB where graduate students, research assistants, and postdoctoral fellows often hold relatively short-term positions\textsuperscript{\protect\hyperlink{ref-D4C4k4ak}{41}}.
Often with these types of short-term research projects, once the contract expires, the research code upkeep tends to fall off as the researchers move on to new projects.
Additionally, if the code is kept on only the researcher's personal devices, it becomes increasingly difficult to access the code for future uses.
As projects or contracts end, there should be a handover period of this code in order for the next cohort of researchers to reuse what was already developed\textsuperscript{\protect\hyperlink{ref-4ny1onB0}{8},\protect\hyperlink{ref-D4C4k4ak}{41}}.

GitHub facilitates project continuity among research code by providing tools that make this handover period easier.
As we have already mentioned, using Git for code in Ecology and Evolution can allow for a ``paper trail'' of sorts to be created for the code, thus allowing for future users of the code access to the entire history of the project\textsuperscript{\protect\hyperlink{ref-3DKwn1sY}{14}}.
Additionally, GitHub allows for repositories and organizations to have designated Code Owners\textsuperscript{\protect\hyperlink{ref-s91uGRZ2}{42}}; these code owners can change through time allowing for the transition of code from one cohort of researchers to the next\textsuperscript{\protect\hyperlink{ref-iIEKCTLU}{30}}.
There are also multiple means by which to archive code dependency information with GitHub.
Dependencies can be described in README files, added to environment.yml and/or an environment.txt file, or stored in a repository using tools such as renv (\url{https://rstudio.github.io/renv/articles/renv.html}).
This ensures that as packages continue to develop with new syntax, code will still run as it did when it was developed.

@@

There have been many calls for researchers outside of the software development community to join the 73 million GitHub users for their collaborative research\textsuperscript{\protect\hyperlink{ref-10ghgV3S8}{3},\protect\hyperlink{ref-1Du6fzB8g}{29},\protect\hyperlink{ref-UsTxAq4f}{43}}.
This call comes in light of both the continual shift toward open-science and increasing computational and data requirements in EEB.
Until now, resources and practical guidance specifically focused on using GitHub within the EEB community have been dispersed in blog posts and video tutorials (\protect\hyperlink{tips}{Box 2} Box\textsuperscript{\protect\hyperlink{ref-box:tips}{\textbf{box:tips?}}}).
We felt these resources have been extremely useful for us to learn to use GitHub in our own work, and that a collation of the main ideas into one medium, while adding on our personal perspectives, would be of use to the EEB community.

@@

A third barrier may come from general reluctance to share data and code publicly, or technical and logistical issues.
The default public nature of GitHub usage can add additional pressure to students and scientists learning to use the platform.
We also note that large file storage is discouraged (and limited) on GitHub but add-ons do exist (e.g.~\url{https://git-lfs.github.com/}) that permit data storage, and increasing integration between platforms (e.g.\textsuperscript{\protect\hyperlink{ref-lx49NGto}{44}}) allows data to be stored away from GitHub and linked dynamically.
We suspect a major additional barrier to EEB researchers is a distinct lack of GitHub help documents for non-English researchers in ecology and evolution, meaning that EEB researchers potentially miss the opportunity to fully understand the importance of version control, reproducibility, and other benefits of GitHub.

@@

\leavevmode\vadjust pre{\hypertarget{ref-MXxgZJ45}{}}%
\CSLLeftMargin{28. }
\CSLRightInline{Dawson, C. \emph{Building tools with GitHub: customize your workflow}. (O'Reilly, 2016).}

\leavevmode\vadjust pre{\hypertarget{ref-1Du6fzB8g}{}}%
\CSLLeftMargin{29. }
\CSLRightInline{Crystal‐Ornelas, R. \emph{et al.} \href{https://doi.org/10.1029/2021ea001797}{A Guide to Using GitHub for Developing and Versioning Data Standards and Reporting Formats}. \emph{Earth Space Sci} \textbf{8}, (2021).}

\leavevmode\vadjust pre{\hypertarget{ref-iIEKCTLU}{}}%
\CSLLeftMargin{30. }
\CSLRightInline{Hampton, S. E. \emph{et al.} \href{https://doi.org/10.1890/es14-00402.1}{The Tao of open science for ecology}. \emph{Ecosphere} \textbf{6}, art120 (2015).}

\leavevmode\vadjust pre{\hypertarget{ref-uBJwnPbq}{}}%
\CSLLeftMargin{31. }
\CSLRightInline{Mislan, K. A. S., Heer, J. M. \& White, E. P. \href{https://doi.org/10.1016/j.tree.2015.11.006}{Elevating The Status of Code in Ecology}. \emph{Trends in Ecology \&amp; Evolution} \textbf{31}, 4--7 (2016).}

\leavevmode\vadjust pre{\hypertarget{ref-1HZdsK5Kn}{}}%
\CSLLeftMargin{32. }
\CSLRightInline{Baker, M. \href{https://doi.org/10.1038/533452a}{1,500 scientists lift the lid on reproducibility}. \emph{Nature} \textbf{533}, 452--454 (2016).}

\leavevmode\vadjust pre{\hypertarget{ref-K7nbP1Ty}{}}%
\CSLLeftMargin{33. }
\CSLRightInline{Crystal-Ornelas, R. \emph{et al.} \emph{Not just for programmers: How GitHub can accelerate collaborative and reproducible research in ecology and evolution}. \url{https://SORTEE-Github-Hackathon.github.io/manuscript/} (2022).}

>>>>

This manuscript
(\href{https://SORTEE-Github-Hackathon.github.io/manuscript/v/dc28d7dd993d0f5ca61b1bd7fc59d1cc22e39ce2/}{permalink})
was automatically generated
from \href{https://github.com/SORTEE-Github-Hackathon/manuscript/tree/dc28d7dd993d0f5ca61b1bd7fc59d1cc22e39ce2}{SORTEE-Github-Hackathon/manuscript@dc28d7d}
on June 23, 2022.

@@

\hypertarget{hosting-a-website}{%
\subsubsection{Hosting a website}\label{hosting-a-website}}

Personal or laboratory websites can improve the sharing of research findings, build online presence, and increase coordination of research efforts\textsuperscript{\protect\hyperlink{ref-HiIPSSHV}{28}}.
Despite many researchers in ecology and evolution having little experience in building or hosting webpages, many tools have been developed to help this process.
Static websites can now be easily built using independent software and languages, such as Jekyll (\url{https://jekyllrb.com}), Hugo (\url{https://gohugo.io}), Quarto (\url{https://quarto.org}), and wowchemy (\url{https://wowchemy.com}), or with the help of dependencies in the programming languages scientists commonly use, such as the \texttt{distill} (\url{https://github.com/rstudio/distill}) and \texttt{blogdown} (\url{https://bookdown.org/yihui/blogdown}) R packages.
The resulting files from static websites can then be hosted in repositories, from which one can activate GitHub's Pages (\url{https://pages.github.com}) feature, allowing for the direct live hosting of HyperText Markup Language (HTML), Cascading Style Sheets (CSS) and JavaScript files within a free github.io domain.
Website templates are readily available on GitHub (https://github.com/topics/website-template) or in user-specific repositories, which can be forked (\protect\hyperlink{definitions}{Box 1}) and customized\textsuperscript{\protect\hyperlink{ref-MXxgZJ45}{29}}.
Aside from free hosting services, GitHub Pages also allows websites to be autogenerated and instantly updated following content modifications\textsuperscript{\protect\hyperlink{ref-RVetqmsg}{7}}.
Creating and hosting websites on GitHub pages can be often time-consuming, as this process usually requires more knowledge in web design than out-of-the-box platforms (\emph{e.g.}, Wix, Weebly, Google Sites).
However, free hosting, widely available template customization, and versioning are strong advantages over the alternatives.

@@

Government, funding agencies, and publishers exercise rigorous open access data policies and mandates\textsuperscript{\protect\hyperlink{ref-PLmDFZrm}{30},\protect\hyperlink{ref-1Hcf13Q0k}{31}}.
However, code and data sharing may be met by individual reluctance, temporary embargoes, or partially prevented by privacy and confidentiality reasons\textsuperscript{\protect\hyperlink{ref-1Ch6LSHef}{32}--\protect\hyperlink{ref-1CzUZwyU2}{34}}.
Still, data deposition and ensuring its availability can amplify the outreach of published studies\textsuperscript{\protect\hyperlink{ref-666HppfO}{35}}, increase citation rates\textsuperscript{\protect\hyperlink{ref-1CcAUn3Lu}{36}}, and among many other reasons, enables the reproducibility and robustness of scientific advances\textsuperscript{\protect\hyperlink{ref-4LaijDIZ}{37}--\protect\hyperlink{ref-1HZdsK5Kn}{39}}.
While public repositories on GitHub make it easy to store and share data files, they are not considered long-term repositories for research materials.
This is because GitHub, a for-profit company, does not have long-term data availability guarantees, allowing users to delete or make repositories private after publication.
Also, GitHub does not issue Digital Object Identifiers (DOI) for content uploaded to their servers.
DOIs are persistent and citable unique alpha-numeric identifiers assigned to digitally stored research materials.
Because of this, scientists sharing code and data through GitHub are strongly encouraged to independently submit their research materials to long-term data archives (\emph{e.g.}, Zenodo, Figshare, Dryad, OSF\textsuperscript{\protect\hyperlink{ref-10ghgV3S8}{3},\protect\hyperlink{ref-kEX5dgzK}{9},\protect\hyperlink{ref-1Du6fzB8g}{40},\protect\hyperlink{ref-tbl:compare}{\textbf{tbl:compare?}}}.
Some of these options (Zenodo, Figshare and OSF) integrate with GitHub, allowing project, code, and data releases (\protect\hyperlink{definitions}{Box 1}) to be archived with versioned, citable DOIs.
Linking GitHub repositories with a DOI helps research become findable, properly cited, and can ensure long-term stability\textsuperscript{\protect\hyperlink{ref-iIEKCTLU}{41}}.

An important aspect of making code and data citable and reusable is to add an appropriate licence to protect intellectual property.
Code published without a licence is under exclusive copyright (by default), protecting it from copy, distribution, and modifications.
One may grant specific rights to their code for reuse by adding licensing files and specifications within GitHub repositories\textsuperscript{\protect\hyperlink{ref-TOsASkn5}{42}}.
The Choose a License (\url{https://choosealicense.com/non-software/}) website offers further guidance on the licenses available for research and creative products.
For example, Creative Commons (CC; \url{https://creativecommons.org/licenses/}) licenses can specify that shared code is intended for a specific analysis.
A CC BY 4.0 license specifies that any code (or other creative products) must be appropriately credited to its original author when distributed, adapted or reused.
A GNU General Public License v3 provides the same attributions above, but requires that all modified or adapted versions to be made public.

\hypertarget{collaborative-and-asynchronous-code-editing}{%
\subsubsection{Collaborative and asynchronous code editing}\label{collaborative-and-asynchronous-code-editing}}

Researchers can asynchronously communicate and individually or collaboratively work in GitHub.
By forking, branching, and cloning repositories (\protect\hyperlink{definitions}{Box 1}), researchers can simultaneously work on different issues of the same research project to then later merge changes to the main branch project with pull requests.
Explicit project organization and increased communication in GitHub Issues, Discussions, or within pull requests can help with project development and with potential merge conflicts due to users simultaneously working on the same sections\textsuperscript{\protect\hyperlink{ref-1BJcvyTmV}{43}}.
Moreover, version control\textsuperscript{\protect\hyperlink{ref-K7nbP1Ty}{44}} and commit reversal features allow researchers to track and progress without worrying about irreparably modifying someone else's work.
By enabling more comprehensive remote collaboration, GitHub encourages the exchange of ideas among researchers at different institutions and in different countries, which can serve to improve the quality of the research itself by providing open access to data and code.
In academic settings, GitHub can also facilitate interactions between research advisors and advisees, providing a platform for students or other trainees to share in-progress code, and flag specific challenges or questions for their supervisors or mentors (\protect\hyperlink{tbl:roles}{Table 1}).
Periodic code review\textsuperscript{\protect\hyperlink{ref-hm9PaCLD}{45}} can also help advisors identify errors early in the process, and inform further training and mentorship to fill gaps in skills.

@@

Beyond supporting collaborative code development, GitHub can be used for writing manuscripts.
Writing a manuscript and storing its associated data and code in GitHub increases scientific reproducibility because text, code, and data can be found in one place.
Writing a manuscript on GitHub may take more time when compared to using conventional text processors\textsuperscript{\protect\hyperlink{ref-4ny1onB0}{8}}.
Nevertheless, GitHub has many features that can allow for a powerful collaborative workflow when writing manuscripts.
Text documents stored and versioned in GitHub can be instantly displayed when written in Markdown, a lightweight markup language increasingly popular among scientists.
Co-authors can contribute changes or suggest revisions to a manuscript written in GitHub through pull requests (\protect\hyperlink{definitions}{Box 1}).
Pull requests provide line-by-line views of proposed changes, which can be commented, modified, or approved by designated reviewers and collaborators (\protect\hyperlink{tbl:roles}{Table 1}).
Relevant literature or issues can be made using the Discussions and Issues features.
Moreover, real-time collaboration on text documents stored in GitHub repositories can be achieved with the help of other platforms (\emph{e.g.}, HackMD (\url{https://hackmd.io} for Markdown documents).

We wrote this manuscript using Manubot\textsuperscript{\protect\hyperlink{ref-YuJbg3zO}{46}}, a workflow implemented in GitHub to automatically render manuscripts and automate bibliographical tasks.
Manubot uses GitHub's automation workflow, GitHub Actions, to combine and convert individual Markdown files into a single LaTeX document, which can then be converted to a Word or PDF document, and displayed as a webpage.
Citations and bibliographic references are automatically managed with citable persistent identifiers (\emph{e.g.}, DOIs, PubMed IDs, ISBNs, URLs).
The resulting manuscript can be rendered with document templates and citation style language formatting to meet journal formatting requirements.
Every change made to the manuscript triggers its rendering, so that updates are readily displayed and made publicly available.
Additional GitHub Actions can be integrated with Manubot, such as ones creating figures or generating tables (\emph{e.g.}, https://github.com/SORTEE-Github-Hackathon/manuscript/tree/main/.github/workflows).
For more examples of manuscripts built via Manubot, see\textsuperscript{\protect\hyperlink{ref-YuJbg3zO}{46}}.

\hypertarget{peer-review}{%
\subsubsection{Peer review}\label{peer-review}}

Peer review is the standard process for assessing and judging whether research done in ecology and evolution should be published in a scientific journal.
GitHub provides an open and transparent platform that can be used for either directly providing feedback on research products or addressing changes recommended by reviewers.
GitHub Issues can be used to organize and discuss reviewer suggestions and to assign them to co-authors (\emph{e.g.}, \url{https://github.com/SORTEE-Github-Hackathon/manuscript/issues?q=label\%3A\%22Reviewer+Comment\%22+}).
When reviewer comments are posted as separate issues, authors can comment on the issues to discuss possible changes and assign themselves to indicate which comments they intend to handle.
Co-authors can then integrate their edits and responses to reviewers using pull requests.

GitHub can also assist reviewers during the peer review process.
If the code associated with a manuscript is made available at the time of submission (\emph{e.g.}, as a link to a GitHub repository within the Data Availability Statement), peer reviewers may be able to offer more comprehensive suggestions on the code and written materials, potentially recognizing errors before publication.
Certain journals or software development communities require submitted work or research code to be hosted on GitHub and their review processes make use of GitHub Issues {[}\emph{e.g.}, rOpenSci (\url{https://ropensci.org/software-review/}); Journal of Open Source Software (\url{https://joss.readthedocs.io/en/latest/submitting.html}){]}.
rOpenSci's efforts have resulted in many well-used R packages for ecology research including \texttt{rfishbase}\textsuperscript{\protect\hyperlink{ref-Xsdcv6q}{47}} and \texttt{taxize}\textsuperscript{\protect\hyperlink{ref-FVBWKkZu}{48}}.

@@

Conventional research practices typically rely on one or two people running and checking the data analyses, while most coauthors (and readers of the subsequent publication) see only the final results and a verbal description of the analytical steps.
In the developmental stages, collaborators can directly see the code for the analysis, manipulate and explore the data themselves, and check for errors.
Cynically, there is also more insurance against nefarious colleagues that may be tempted to distort results\textsuperscript{\protect\hyperlink{ref-18PTmKJkq}{49}}.
Collaborators as well as pre-publication and post-publication reviewers are better positioned to discover questionable findings if they have full and transparent access to the project.

This transparency can similarly be extended beyond coauthors to the entire scientific community.
Publishing the data and reproducible workflows along with the manuscript allows any reader to review the analysis and reproduce the experiment\textsuperscript{\protect\hyperlink{ref-NOgBWVAr}{50}}.
Supplying code for (novel) methods that are proposed or used also reduces barriers to knowledge and can greatly improve the ability of others to build on existing work, resulting in greater proliferation and accessibility for a broader audience.
GitHub even provides a useful Discussions Forum (\url{https://docs.github.com/en/discussions}) that aids the direct communication with repository owners, as well as the GitHub Community (\url{https://github.community/}) forum for more general questions and sharing of expertise (\protect\hyperlink{tbl:roles}{Table 1}.

@@

The development of research code continues, and so does the need to consider project continuity, especially in EEB where graduate students, research assistants, and postdoctoral fellows often hold relatively short-term positions\textsuperscript{\protect\hyperlink{ref-D4C4k4ak}{51}}.
Often with these types of short-term research projects, once the contract expires, the research code upkeep tends to fall off as the researchers move on to new projects.
Additionally, if the code is kept on only the researcher's personal devices, it becomes increasingly difficult to access the code for future uses.
As projects or contracts end, there should be a handover period of this code in order for the next cohort of researchers to reuse what was already developed\textsuperscript{\protect\hyperlink{ref-4ny1onB0}{8},\protect\hyperlink{ref-D4C4k4ak}{51}}.

GitHub facilitates project continuity among research code by providing tools that make this handover period easier.
As we have already mentioned, using Git for code in Ecology and Evolution can allow for a ``paper trail'' of sorts to be created for the code, thus allowing for future users of the code access to the entire history of the project\textsuperscript{\protect\hyperlink{ref-3DKwn1sY}{14}}.
Additionally, GitHub allows for repositories and organizations to have designated Code Owners\textsuperscript{\protect\hyperlink{ref-s91uGRZ2}{52}}; these code owners can change through time allowing for the transition of code from one cohort of researchers to the next\textsuperscript{\protect\hyperlink{ref-iIEKCTLU}{41}}.
There are also multiple means by which to archive code dependency information with GitHub.
Dependencies can be described in README files, added to environment.yml and/or an environment.txt file, or stored in a repository using tools such as renv (\url{https://rstudio.github.io/renv/articles/renv.html}).
This ensures that as packages continue to develop with new syntax, code will still run as it did when it was developed.

@@

There have been many calls for researchers outside of the software development community to join the 73 million GitHub users for their collaborative research\textsuperscript{\protect\hyperlink{ref-10ghgV3S8}{3},\protect\hyperlink{ref-1Du6fzB8g}{40},\protect\hyperlink{ref-UsTxAq4f}{53}}.
This call comes in light of both the continual shift toward open-science and increasing computational and data requirements in EEB.
Until now, resources and practical guidance specifically focused on using GitHub within the EEB community have been dispersed in blog posts and video tutorials (\protect\hyperlink{tips}{Box 2} Box\textsuperscript{\protect\hyperlink{ref-box:tips}{\textbf{box:tips?}}}).
We felt these resources have been extremely useful for us to learn to use GitHub in our own work, and that a collation of the main ideas into one medium, while adding on our personal perspectives, would be of use to the EEB community.

@@

A third barrier may come from general reluctance to share data and code publicly, or technical and logistical issues.
The default public nature of GitHub usage can add additional pressure to students and scientists learning to use the platform.
We also note that large file storage is discouraged (and limited) on GitHub but add-ons do exist (e.g.~\url{https://git-lfs.github.com/}) that permit data storage, and increasing integration between platforms (e.g.\textsuperscript{\protect\hyperlink{ref-lx49NGto}{54}}) allows data to be stored away from GitHub and linked dynamically.
We suspect a major additional barrier to EEB researchers is a distinct lack of GitHub help documents for non-English researchers in ecology and evolution, meaning that EEB researchers potentially miss the opportunity to fully understand the importance of version control, reproducibility, and other benefits of GitHub.

@@

\leavevmode\vadjust pre{\hypertarget{ref-HiIPSSHV}{}}%
\CSLLeftMargin{28. }
\CSLRightInline{Smaglik, P. \href{https://doi.org/10.1038/nj7142-347a}{Creating better lab websites gives potential collaborators and recruiters a clearer window into your world.} \emph{Nature} \textbf{447}, 347--347 (2007).}

\leavevmode\vadjust pre{\hypertarget{ref-MXxgZJ45}{}}%
\CSLLeftMargin{29. }
\CSLRightInline{Dawson, C. \emph{Building tools with GitHub: customize your workflow}. (O'Reilly, 2016).}

\leavevmode\vadjust pre{\hypertarget{ref-PLmDFZrm}{}}%
\CSLLeftMargin{30. }
\CSLRightInline{Tenopir, C. \emph{et al.} \href{https://doi.org/10.1371/journal.pone.0229003}{Data sharing, management, use, and reuse: Practices and perceptions of scientists worldwide}. \emph{PLoS ONE} \textbf{15}, e0229003 (2020).}

\leavevmode\vadjust pre{\hypertarget{ref-1Hcf13Q0k}{}}%
\CSLLeftMargin{31. }
\CSLRightInline{Nugroho, R. P., Zuiderwijk, A., Janssen, M. \& de Jong, M. \href{https://doi.org/10.1108/tg-03-2014-0008}{A comparison of national open data policies: lessons learned}. \emph{Transforming Government: People, Process and Policy} \textbf{9}, 286--308 (2015).}

\leavevmode\vadjust pre{\hypertarget{ref-1Ch6LSHef}{}}%
\CSLLeftMargin{32. }
\CSLRightInline{Wicherts, J. M., Bakker, M. \& Molenaar, D. \href{https://doi.org/10.1371/journal.pone.0026828}{Willingness to Share Research Data Is Related to the Strength of the Evidence and the Quality of Reporting of Statistical Results}. \emph{PLoS ONE} \textbf{6}, e26828 (2011).}

\leavevmode\vadjust pre{\hypertarget{ref-SLq38RVv}{}}%
\CSLLeftMargin{33. }
\CSLRightInline{Figueiredo, A. S. \href{https://doi.org/10.3389/fpubh.2017.00327}{Data Sharing: Convert Challenges into Opportunities}. \emph{Front. Public Health} \textbf{5}, (2017).}

\leavevmode\vadjust pre{\hypertarget{ref-1CzUZwyU2}{}}%
\CSLLeftMargin{34. }
\CSLRightInline{Tenopir, C. \emph{et al.} \href{https://doi.org/10.1371/journal.pone.0134826}{Changes in Data Sharing and Data Reuse Practices and Perceptions among Scientists Worldwide}. \emph{PLoS ONE} \textbf{10}, e0134826 (2015).}
